DOpElib:\\ Differential Equations and Optimization Environment

### 2016-11-17 Examples/Documentation/examples.tex

\item We now have to change the target in the \texttt{Makefile}, 
  otherwise our new example will refer to the executable of the old one.

  Thus we open the \texttt{Makefile} in the directory \texttt{MyWonderfulFirstExample}
  to change the name of the executable. In our example you will find a line
\begin{verbatim}
target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
           (dope_dimension)d-$(deal_II_dimension)d
\end{verbatim}
Here we need to replace the target to a new name to avoid any 
possible complication. Thus for example we change the line to 
\begin{verbatim}
target   = $(BINDIR)/MyWonderfulFirstExample
\end{verbatim}
This will build the executable file in the standard binary directory 
of \texttt{DOpElib}, i.e., \texttt{\$(DOpE)/bin}, under the name 
\texttt{MyWonderfulFirstExample}.

\textit{Note that you should use a different name for all your 
examples as otherwise you may experience a lot of strange behavior, as
the executable will be overwritten by all your examples!}

@@

  Change to the \texttt{Test} sub-directory. And then modify the
  test-script to contain the new name of the executable.
  Assuming you want to use Emacs, open the file \texttt{test.sh}
\begin{verbatim}
PDE/StatPDE/Example1/Test> emacs test.sh
\end{verbatim}
where, in our example you find the line
\begin{verbatim}
PROGRAM=../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d
\end{verbatim} 
if you made a copy of an other example the part \texttt{DOpE-PDE-StatPDE-Example1-2d-2d}
may differ. These lines need to be replaced with the new name of the 
executable, i.e., for our given example
\begin{verbatim}
PROGRAM=../../../../../bin/MyWonderfulFirstExample
\end{verbatim}

\item  Now, you are prepared to change any of the problem
  dependent data in information in the files 
\begin{verbatim}
main.cc, localpde.h, functionals.h, localfunctional.h, etc
\end{verbatim}

>>>>

\item You will find a file \texttt{CMakeLists.txt} in the folder 
\texttt{MyWonderfulFirstExample}. Open this file, in it you can 
find the line 
\begin{verbatim}
  SET(TARGET "DOpE-PDE-StatPDE-Example1")
\end{verbatim}
the string \texttt{DOpE-PDE-StatPDE-Example1} will be the name of the
executable of your program. Change it to something suitable, e.g.,
\begin{verbatim}
  SET(TARGET "MyWonderfulFirstExample")
\end{verbatim}
Moreover, this file contains the lines 
\begin{verbatim}
SET(dope_dimension 2)
SET(deal_dimension 2)
\end{verbatim}
which define the dimension of the domain for the control-variable
(\texttt{dope\_dimension})
and the PDE solution (\texttt{deal\_dimension}). If for your example 
one of these differs from $2$ adjust the number accordingly.
This file will not need any further modifications.

\item  Now, you are prepared to change any of the problem
  dependent data in information in the files 
\begin{verbatim}
main.cc, localpde.h, functionals.h, localfunctional.h, etc
\end{verbatim} 
If above you have changed the dimensions, make sure to adjust all
files accordingly!

\item The cmake system can - in principle be used
  ''in-source'' to create the executable. 
  However, this may break some of the automated tests (later),
  so you are encouraged to proceed differently. To do this end
  proceed as follows in the directory \texttt{MyWonderfulFirstExample}:
  \begin{verbatim}
MyWonderfulFirstExample$ mkdir build
MyWonderfulFirstExample$ cd build
MyWonderfulFirstExample/build$ cmake -DCMAKE_BUILD_TYPE=Release ../
\end{verbatim}
which will configure the build (if you want to debug it is useful to
replace the string \texttt{Release} with \texttt{Debug}). 
Once this is done, you can compile and run the code:
\begin{verbatim}
MyWonderfulFirstExample/build$ make 
MyWonderfulFirstExample/build$ cd ..
MyWonderfulFirstExample$ ./MyWonderfulFirstExample
\end{verbatim}
(Assuming that no errors occurred during the make call)

\item The \texttt{Makefile} in the directory is present only to
  preserve backward compatibility. If you wish to use the automated
  build/test routines in DOpE lib, you need to make sure that it its
  configured correctly. If you just want to use the cmake capabilities 
  you can safely ignore this passage.

  Furthermore, if you have followed the instructions above, then no
  changes will be needed.

  If you have moved the directory to some other place, i.e., it is not
  in the same folder as Example1, then open the \texttt{Makefile} 
  in the directory \texttt{MyWonderfulFirstExample}
  you will find the line 
  \begin{verbatim}
  DOpE = ../../../../
  \end{verbatim}
  which points to the root directory of your DOpE installation.
  Adjust the path to match accordingly.

@@

  Change to the \texttt{Test} sub-directory. And then modify the
  test-script to contain the new name of the executable.
  Assuming you want to use Emacs, open the file \texttt{test.sh}
\begin{verbatim}
PDE/StatPDE/Example1/Test> emacs test.sh
\end{verbatim}
  where, in our example you find the line
\begin{verbatim}
PROGRAM=../DOpE-PDE-StatPDE-Example1-2d-2d
\end{verbatim} 
  if you made a copy of an other example the part \texttt{DOpE-PDE-StatPDE-Example1-2d-2d}
  may differ. These lines need to be replaced with the new name of the 
  executable, i.e., for our given example
\begin{verbatim}
PROGRAM=../MyWonderfulFirstExample
\end{verbatim}

### 2016-11-17 Examples/Documentation/introduction.tex

\section{License information}
Copyright (C) 2012--2014 by the DOpElib authors\\[2mm]
%
This file is part of DOpElib\\[2mm]
%
DOpElib is free software: you can redistribute it
and/or modify it under the terms of the GNU General Public
License as published by the Free Software Foundation, either
version 3 of the License, or (at your option) any later
version.\\[2mm]
%
DOpElib is distributed in the hope that it will be
useful, but WITHOUT ANY WARRANTY; without even the implied
warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
PURPOSE.  See the GNU General Public License for more
details.\\[2mm]
%
Please refer to the file LICENSE.TXT included in this distribution
for further information on this license.

@@

\section{Contributors \& Developers}
\label{sec:contrib}
The library is currently maintained by 
\begin{itemize}
  \item Christian Goll (Heidelberg University)
  \item Thomas Wick (Johann Radon Institute for Computational and Applied Mathematics (RICAM))
  \item Winnifried Wollner (Technische Universität Darmstadt)
\end{itemize}

@@

@@

\subsubsection{SNOPT}
If you would like to use the features offered in our SNOPT wrapper. You will 
need to obtain a license for \texttt{SNOPT} 
\url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
Unfortunately this is at present not available for free, but you should 
check if 
there is a department license already available.
For further information you should consult the file 
\texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
\texttt{deal.II} with at least the following options:\\
\texttt{./configure --with-umfpack --with-blas --with-lapack} (up to 7.3.0)
or enabling UMFPACK in the ccmake options from version 8.0.0 on.

@@

\section{Installation}\label{sec_installation}
\subsection{After Version 2.0 - CMake build system}
To work with DOpElib, your need necessarily deal.II, which installation 
we describe first. Afterwards, the DOpElib installation is described and finally
other optional packages might be installed.
\begin{itemize}
\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
\texttt{\textasciitilde/deal.II}. \\[1mm]
If you prefer another position you can install it anywhere but need 
to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
will be found by cmake, i.e.,
\begin{lstlisting}
export DEAL_II_DIR=[path to deal.ii]
\end{lstlisting}
Detailed installing instructions for deal.II (here the last version
8.4.0) can be found on 
\texttt{http://www.dealii.org/8.4.0/readme.html}. The deal.II 
installation instructions are descriptive enough and we omit 
any further comments and refer to their webpage.
%

>>>>

\section{License information}
Copyright (C) 2012--2016 by the DOpElib authors\\[2mm]
%
This file is part of DOpElib\\[2mm]
%
DOpElib is free software: you can redistribute it
and/or modify it under the terms of the GNU General Public
License as published by the Free Software Foundation, either
version 3 of the License, or (at your option) any later
version.\\[2mm]
%
DOpElib is distributed in the hope that it will be
useful, but WITHOUT ANY WARRANTY; without even the implied
warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
PURPOSE.  See the GNU General Public License for more
details.\\[2mm]
%
Please refer to the file LICENSE.TXT included in this distribution
for further information on this license.

@@

\section{Contributors \& Developers}
\label{sec:contrib}
The library is currently maintained by 
\begin{itemize}
  \item Thomas Wick (Centre de Mathématiques Appliquées (CMAP) École Polytechnique)
  \item Winnifried Wollner (Technische Universität Darmstadt)
\end{itemize}

@@

for the use of MPI use

\begin{lstlisting}
-DDEAL_II_WITH_MPI=true
\end{lstlisting}

and for the use of distributed meshes over several computers the p4est
library is needed in deal.II to use this use the flags (in addition to the use of MPI)
\begin{lstlisting}
-DDEAL_II_WITH_P4EST=true -DP4EST_DIR=[path to p4est]
\end{lstlisting}

@@

\subsubsection{SNOPT}
If you would like to use the features offered in our SNOPT wrapper. You will 
need to obtain a license for \texttt{SNOPT} 
\url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
Unfortunately this is at present not available for free, but you should 
check if 
there is a department license already available.
For further information you should consult the file 
\texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
\texttt{deal.II} with at least with enabled UMFPACK.

@@

\section{Installation}\label{sec_installation}
To work with DOpElib, your need necessarily deal.II, which installation 
we describe first. Afterwards, the DOpElib installation is described and finally
other optional packages might be installed.
\begin{itemize}
\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
\texttt{\textasciitilde/deal.II}. \\[1mm]
If you prefer another position you can install it anywhere but need 
to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
will be found by cmake, i.e.,
\begin{lstlisting}
export DEAL_II_DIR=[path to deal.ii]
\end{lstlisting}
Detailed installing instructions for deal.II (here the last version
8.4.0) can be found on 
\texttt{http://www.dealii.org/8.4.0/readme.html}. The deal.II 
installation instructions are descriptive enough and we omit 
any further comments and refer to their webpage.
%

### 2016-11-17 Examples/description.tex
Added Nonstationary Example (isothermal Euler) with dG discretization in time + Bugfixes
Further work on optimization with functionals not given by a single integration. Calculation of such functionals is now working, derivative calculation still missing
Corrected umlaute in documentation

\documentclass[a4paper,cleardoubleempty]{scrreprt}
\usepackage{amsmath,amssymb, amsthm, commath}
\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
\usepackage{booktabs}
\usepackage{makeidx}
\usepackage{url}
%\usepackage{sfmath}
\usepackage{color}
\usepackage{graphicx} 
\usepackage{enumerate}
\usepackage{listings}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{DOpElib:\\ 
Differential Equations and Optimization Environment}
\author{
  Thomas Wick\thanks{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
    {thomas.wick@ricam.oeaw.ac.at}}, 
  Winnifried Wollner\thanks{Technische Universität Darmstadt, 
    {wollner@mathematik.tu-darmstadt.de}}
}

@@

@@

@@

>>>>

\documentclass[a4paper,cleardoubleempty]{scrreprt}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb, amsthm, commath}
\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
\usepackage{booktabs}
\usepackage{makeidx}
\usepackage{url}
%\usepackage{sfmath}
\usepackage{color}
\usepackage{graphicx} 
\usepackage{enumerate}
\usepackage{listings}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{DOpElib:\\ 
Differential Equations and Optimization Environment}
\author{
  Thomas Wick\thanks{École Polytechnique, 
    {thomas.wick@polytechnique.edu}}, 
  Winnifried Wollner\thanks{Technische Universität Darmstadt, 
    {wollner@mathematik.tu-darmstadt.de}}
}

@@

\subsection{Isothermal Euler equations}
\label{PDE_isothermal_euler}
\input{PDE/InstatPDE/Example7/content.tex}

@@

\subsection{Parameter control with a nonlinear PDE from fluid dynamics - with cost functional
not given as an integral but as function of an integral}
\label{OPT_Stat_Param_Nonlin_Fluid_extension}
\input{OPT/StatPDE/Example10/content.tex}

@@

\subsection{Minimizing the spatial mean-value - cost functional involves functions of integrals}
\label{OPT_Instat initial-value extension}
\input{OPT/InstatPDE/Example4/content.tex}
